Hippocampal replay of experience at real-world speeds

### 2020-10-20 Article.tex
Biorvix Draft 1

%TC:break Abstract
%the command above serves to have a word count for the abstract
\begin{abstract}
Representations of past and possible future experiences play a critical role in memory and decision-making processes. The hippocampus expresses these types of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at \texttildelow20x real-world speeds. Efforts to understand replay typically make multiple assumptions about which events to examine and what sorts of representations constitute replay. We therefore lack a clear understanding of both the prevalence and the range of representational dynamics associated with replay. Here we develop a state space model that uses a combination of movement dynamics of different speeds to capture the spatial content and time evolution of replay during SWRs. We find that the large majority of replay events contain spatially coherent, interpretable content. Furthermore, most events progress at real-world, rather than accelerated, movement speeds, consistent with actual experiences. 
\end {abstract}
%TC:break main
%the command above serves to have a word count for the abstract

@@

An ideal approach to identifying and quantifying the dynamics of generative representations would circumvent these problems. It would use all of the available spiking data to yield the most accurate decoded positions. It would be applicable to either thresholded events or to all of the data to permit an unbiased assessment of the prevalence and nature of generative activity. It would use very small temporal bins (1 or 2 ms) to allow for very rapid representational movement and provide information about the certainty of the decoded estimates. It would be able to capture a range of movement dynamics -- how the position representation evolves over time. An ideal model would capture key dynamics of interest including stationary or unchanging representations, trajectories that progress through space at constant or variable speeds, and disorganized, spatially incoherent events. It would provide a robust statistical assessment of confidence for each dynamic. Finally, where assumptions are made, it would provide well defined parameters whose values can be explored systematically to understand their influence on the results.

@@

Next, we specify movement dynamic models for how latent position---the "mental" position of the animal represented by the cells---evolves over time for a given dynamic (Figure \ref{1}C). We do this by defining a state transition matrix that defines how the latent position can move from the previous time step. Previous findings suggest that replay may exhibit at least three distinct types of movement dynamics: stationary \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, continuous \cite{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \cite{PfeifferAutoassociativedynamicsgeneration2015}. We therefore define state transition models to capture each of these movement dynamics.

In the stationary movement dynamic, the latent position does not change between time steps. The state transition matrix can thus be defined as an identity matrix, which predicts the next position will be the same as the last position (subject to the size of the position bin). In the continuous movement dynamic, the latent position is most likely to be  "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix. This means that, for a given latent position, the probability of moving to another position is modeled by a Gaussian centered at that position and "spatially close" is defined by the variance of the Gaussian. In our case, since replay has been shown to move at speeds much faster than the animal's speed \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015}, we set the variance of the Gaussian to 6.0 cm. This ensures that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (or equivalently, this means that latent speeds from 0 to \texttildelow25 m/s are most likely).  Finally, in the fragmented movement dynamic, the latent position can move to any available position instantaneously. We model this using a uniform state transition matrix, which makes transitions to each position equally likely.

Finally, we specify how likely each movement dynamic is to persist in time versus changing to another dynamic (Figure \ref{1}D) via a state transition matrix between dynamics. In order to be conservative, we assume that each movement dynamic is likely to dominate for \texttildelow100 ms, which is approximately the duration of a SWR event. There is, however, a small probability of switching to one of the other movement dynamics. Accordingly, we set the probability of staying in a dynamic to 0.98 for each 2 ms time step, which corresponds to an expected duration of 100 ms for staying in a particular dynamic (the Markov assumption of the model means that the probability of staying in a dynamic follows a geometric distribution). Importantly, the data drives the estimated dynamic, so even if the probability of a particular movement dynamic is high, data providing clear evidence for a change in dynamic can, within the time frame of \texttildelow10 ms, drive a change to a different estimated dynamic, as in our simulated example. In line with this, note that below we show that our results are relatively insensitive to the value of this parameter.

@@

\begin{figure*}%[tbhp]
\centering
\includegraphics[width=0.80\linewidth]{figures/Figure2/Figure2_v5}
\caption{The model can decode hippocampal replay trajectories using either sorted and clusterless spikes from the same SWR event. \textbf{(A)} Decoding using sorted spikes. The top panel shows 31 cells on a W-track ordered according to linearized position by their place field peaks. The middle panel shows the probability of each dynamic over time as in Figure \ref{1}F. Shaded regions correspond to the speed classifications as in Figure \ref{1}G. The bottom panel shows the estimated probability of latent position over the course of the SWR as it travels down the center arm toward the right arm. L, R, C correspond to the position of the left, right and center reward wells respectively. The animal's actual position is indicated by the the magenta dashed line. Underneath is the maximum of the 1D decoded position (the most probable position) projected back onto the 2D track for the sorted decoding. Color indicates time. The animal's actual position is denoted by the pink dot. Light grey lines show the animal's 2D position over the entire recording session. \textbf{(B)} Decoding using clusterless spikes. The top panel shows multiunit spiking activity from each tetrode. Other panels have the same convention as (A). Underneath is the maximum of the 1D decoded position (the most probable position) projected back into 2D using the clusterless decoding.
}
\label{2}
\end{figure*}
\subsection*{Identification of continuous replays with sorted and clusterless spikes}

@@

We next sought to validate the model on real hippocampal data recorded from awake, behaving rats. To ensure that we could capture rapid trajectories many times the speed of the animal, as described by most studies of hippocampal replay, we first decoded a single SWR with sequential population activity (Figure \ref{2}A, top panel) from 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on a W-shaped track. As expected, we observed that the probability of being in the continuous dynamic is high throughout this event, but the probability of being in a stationary dynamic was also noticeable at the beginning and end of the SWR (Figure \ref{2}A, middle panel). Using our speed classification scheme, this means that the speed of the replay starts slower---as a mixture of continuous and stationary dynamics---and then speeds up and slows down again. This is also evident in the posterior probability of linear position over time. This shows that the replay most likely travels down the center arm and up the right arm (Figure \ref{2}A, bottom panel). We can also see this when we project the maximum of the posterior of this trajectory (the most probable "mental" position) to 2D to better see the spatial trajectory on the maze (Figure \ref{2}C). Importantly, when we apply the same model using the 2D position of the animal, we get a similar result as when we use the 1D linearized position (Figure \ref{fig:Figure2-Figure supplement 1}A).

One of our criteria for a more optimal method is for it to use all of the available spiking data. Using only clustered spikes discards any spike events that cannot be uniquely assigned to a putative single neuron, substantially reducing the amount of information that the resultant decoding can use. Additionally, spike sorting is not necessary to recover the underlying neural population dynamics \cite{TrautmannAccurateEstimationNeural2019}, is time-consuming, and often requires inclusion decisions that can vary widely across experimenters. We therefore adopted a "clusterless" approach which directly relates spikes and their multiunit spike waveform features to decode position without spike sorting (see Methods). Clusterless decoding has previously been used to successfully identify theta sequences and replay sequences in the hippocampus \cite{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}. Applying a clusterless decoder to the same SWR event, we get similar classification of the sequence (Figure \ref{2}B, D), both with 1D linearized position and 2D position (Figure \ref{fig:Figure2-Figure supplement 1}B). As predicted, the spatial extent of the event is longer and the estimate of the posterior probability of position is narrower for the clusterless model. This reflects the clusterless model's access to a larger pool of data that provides more information about the extent of the event and more certainty in the latent position and the dynamic (Figure \ref{2}D vs C).

@@

As expected, and providing further support for the validity of the method, we observe many replays that are classified as continuous throughout the entirety of the SWR, similar to those seen using the standard decoder (Figure \ref{fig:Figure2-Figure supplement 2}A-F). However, we also observe many events with spatially coherent content that do not have this structure. For example, we observe trajectories that start in one direction and reverse back to the original position (Figure \ref{3}A, Figure \ref{fig:Figure3-Figure supplement 1}B), trajectories that remain fixed in one position (Figure \ref{3}B, Figure \ref{fig:Figure3-Figure supplement 1}G), and trajectories that jump between arms and between dynamics (Figure \ref{3}C, D, F, Figure \ref{fig:Figure3-Figure supplement 1}F, H, I). We also observe SWRs that are spatially incoherent throughout the SWR (Figure \ref{fig:Figure3-Figure supplement 1}A, D).

@@

Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content \cite{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}. We therefore asked what fraction of classified events contain spatially coherent content, which we define as SWRs containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods). We find that 96\% (22,478 of 23,382) of classified SWRs and 85\% of all SWRs include spatially coherent structure and that this prevalence of spatially coherent structure is consistent across animals (Figure \ref{3}G). We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented or fragmented-continuous mixture dynamics. We find that only 14\% (3,353 of 23,382) of classified SWRs have time points with spatially incoherent structure (Figure \ref{3}G).

@@

Because our model is specified in the context of a latent position associated with different movement dynamics, it allows us to not only classify events in terms of their dynamics, but also to quantify the model's certainty in each position estimate at each moment in time given the model parameters. To do so, we can compute the cumulative spatial size of the 95\% highest posterior density (HPD) region of the latent position estimate. Larger values of the HPD size indicate the model is less certain about position, because the most probable decoded positions are distributed over more of the track at a given time point. In contrast, lower values indicate that the model is more certain about the estimate of position because the extent of the HPD is more concentrated and covers less of the track. Thus, the HPD provides a complementary measure of spatial coherence, and evaluating it allows us to verify that the events we defined as spatially coherent also correspond to events where there is high certainty around the position estimates.

We find that spatially coherent events indeed have lower HPD sizes than events with fragmented dynamics. Figure \ref{4}A and D show two example SWRs that are classified as having stationary and continuous dynamics, respectively. The most probable positions (as estimated by the model) at each time step in these SWRs is concentrated in a small portion of the track and correspondingly, the HPD size is small throughout the SWRs. In contrast, Figure \ref{4}C shows a SWR where the dynamics are fragmented and correspondingly, the most probable positions are more spatially diffuse and the HPD size is much larger. The HPD size also provides insights into the moment-by-moment structure of each event, which can change over the time course of a SWR. An example of this change is shown in Figure \ref{4}B, where the HPD size is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation. Overall, when we examine the average HPD size for each SWR, grouped by dynamic, we find a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E). For the spatially coherent dynamics, the average HPD size for each SWR was much smaller than the spatially incoherent dynamics (median 24 cm, spatially coherent vs. median 238 cm, spatially incoherent, p=2.2e-16, one-sided Mann-Whitney-U). Additionally, the HPD size for the unclassified SWRs was similarly large. We note here that while the size of the HPD will be influenced by the dynamic estimated from the data, it remains possible for continuous or stationary dynamics to correspond to a large HPD region (see outliers in Figure \ref{4}E), indicating less spatial certainty for those specific events. In general, if the data does not exhibit strong place specific firing, the HPD size will be large, regardless of dynamic classification. To show this, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model. The shuffled decodes have much larger HPD size than the real data (epoch \#1: 55 cm vs. 229 cm, p=0.02, real vs. shuffled; epoch \#2: 72 cm vs. 231 cm, p=0.02, real vs. shuffled; one-sided Mann-Whitney-U, \ref{fig:Figure4-Figure supplement 1}).

\begin{figure*}%[tb2]
\centering
\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v6}
\caption{Validation of classification using the 95\% Highest Posterior Density. \textbf{(A-D)}  Examples of the 95 \% Highest Posterior Density. In each column: top panel: Probability of dynamic over time. Shading and labels indicate dynamic categories. Middle panel: Posterior distribution of estimated linear position over time. Magenta dashed line indicates animal's position. Bottom panel: HPD size---the integral of the position bins with the top 95\% probable values. \textbf{(E)} Average 95\% HPD size for each dynamic category. Median 24 cm for spatially coherent vs. median 238 cm for spatially incoherent, p=2.2e-16, one-sided Mann-Whitney-U}
\label{4}
\end{figure*}

\subsection*{Many replay events are slow or stationary}
\begin{figure*}%[tbhp]
\centering
\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v4}
\caption{
Prevalence of classifications. \textbf{(A)} UpSet plot \cite{LexUpSetVisualizationIntersecting2014}---which is similar to a Venn diagram with more than 3 sets---of the most common sets of classifications within each SWR. Each row represents a classification and each column represents a set of classifications, where filled-in black dots with an edge between the dots indicates that multiple classifications are present in the SWR (at temporally distinct times). The sets are ordered by how often they occur as indicated by the bar plot above each category. The total percentage of each classification is indicated by the rotated bar plot on the left. \textbf{(B)} Duration of each dynamic category within a SWR. The box shows the interquartile range (25-75\%) of the data and the whiskers are 1.5 times the interquartile range. Points outside of this are labeled as diamonds. \textbf{(C)} Average distance between the latent position and the animal's actual position for each classification within the SWR. \textbf{(D)} Average speed of the classification within the SWR (excluding classifications with durations < 20 ms). Note that these speeds are calculated using the most probable position (MAP estimate) which can be noisy when the probability of position is flat or multimodal.\textbf{(E)} Kernel density estimate of the position of stationary trajectories on the W-track at least 30 cm away from the animal's position. The shaded region represents the density estimate while the black ticks represent the observed non-local stationary positions. \textbf{(F)} Average tetrode multiunit spike rates for each dynamic category within each SWR (excluding classifications < 20 ms).
}
\label{5}
\end{figure*}

@@

Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of <1 m/s. The next most common category were stationary events \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events ---continuous, extended trajectories through space---made up only about 19\% of classified events. Importantly, events with slower dynamics tended to persist for approximately the same length of time as events (\texttildelow50-100 ms) with faster continuous dynamics, consistent with a brief activation of a representation of a location or a short trajectory through space. Events with slower dynamics could also represent either the animal's current location or locations far from the animal, and similarly, previous work has shown that events with continuous dynamics most often begin at the animal's current location \cite{DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.

@@

\subsection*{Highest Posterior Density}
The 95\% highest posterior density (HPD) is a measurement of the spread of the posterior probability and is defined as the region of the posterior that contains the top 95\% of values of the posterior probability \cite{CasellaStatisticalinference2001}. In this manuscript, we use the HPD to evaluate the uncertainty of the posterior probability of position. We calculate the 95\% HPD by determining the maximum threshold value $h$ that fulfills the follow equality:
$$
\int_{\{x: p(x_{k} \mid O_{1:T}) > h\}} p(x_{k} \mid O_{1:T})dx = 0.95
$$
The 95\% HPD is the set of position bins with posterior values greater than the threshold $\{x : p(x_{k} \mid O_{1:T}) > h\}$. The HPD size is calculated by taking the integral of the members of this set:
$$
\int_{\{x: p(x_{k} \mid O_{1:T}) > h\}} \mathbb{1}dx
$$
which we approximate with a Riemann sum.

>>>>

%TC:break Abstract
%the command above serves to have a word count for the abstract
\begin{abstract}
Representations of past and possible future experiences play a critical role in memory and decision-making processes. The hippocampus expresses these types of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at \texttildelow20x real-world speeds. Efforts to understand replay typically make multiple assumptions about which events to examine and what sorts of representations constitute replay. We therefore lack a clear understanding of both the prevalence and the range of representational dynamics associated with replay. Here we develop a state space model that uses a combination of movement dynamics of different speeds to capture the spatial content and time evolution of replay during SWRs. Using this model, we find that the large majority of replay events contain spatially coherent, interpretable content. Furthermore, most events progress at real-world, rather than accelerated, movement speeds, consistent with actual experiences. 
\end {abstract}
%TC:break main
%the command above serves to have a word count for the abstract

@@

An ideal approach to identifying and quantifying the dynamics of generative representations would circumvent these problems. It would use all of the available spiking data to yield the most accurate decoded positions. It would be applicable to either thresholded events or to all of the data to permit an unbiased assessment of the prevalence and nature of generative activity. It would use very small temporal bins (1 or 2 ms) to allow for very rapid representational movement and provide information about the certainty of the decoded estimates. It would be able to capture a range of movement dynamics from stationary or unchanging representations to trajectories that progress through space at constant or variable speeds to disorganized, spatially incoherent events. It would provide a robust statistical assessment of confidence for each dynamic. Finally, where assumptions are made, it would provide well defined parameters whose values can be explored systematically to understand their influence on the results.

@@

Next, we specify movement dynamic models for how latent position---the "mental" position of the animal represented by the cells---evolves over time for a given dynamic (Figure \ref{1}C). We do this by defining a state transition matrix that defines how the latent position can move from the previous time step (in our case, 2 ms). Previous findings suggest that replay may exhibit at least three distinct types of movement dynamics: stationary \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, continuous \cite{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \cite{PfeifferAutoassociativedynamicsgeneration2015}. We therefore define state transition models to capture each of these movement dynamics.

In the stationary movement dynamic, the latent position does not change between time steps. The state transition matrix can thus be defined as an identity matrix, which predicts the next position will be the same as the last position (subject to the resolution of the position bin). In the continuous movement dynamic, the latent position is most likely to be "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix. This means that, for a given latent position, the probability of moving to another position is modeled by a Gaussian centered at that position and "spatially close" is defined by the variance of the Gaussian. In our case, since replay has been shown to move at speeds much faster than the animal's speed \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015}, we set the variance of the Gaussian to 6.0 cm. This ensures that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (or equivalently, this means that latent speeds from 0 to \texttildelow25 m/s are most likely).  Finally, in the fragmented movement dynamic, the latent position can move to any available position instantaneously. We model this using a uniform state transition matrix, which makes transitions to each position equally likely.

Finally, we specify how likely each movement dynamic is to persist in time versus change to another dynamic (Figure \ref{1}D) via a state transition matrix between dynamics. In order to be conservative, we assume that each movement dynamic is likely to dominate for \texttildelow100 ms, which is approximately the duration of a SWR event. There is, however, a small probability of switching to one of the other movement dynamics. Accordingly, we set the probability of staying in a dynamic to 0.98 for each 2 ms time step, which corresponds to an expected duration of 100 ms for staying in a particular dynamic (the Markov assumption of the model means that the probability of staying in a dynamic follows a geometric distribution). Importantly, the data drives the estimated dynamic, so even if the probability of a particular movement dynamic is high, data providing clear evidence for a change in dynamic can, within the time frame of \texttildelow10 ms, drive a change to a different estimated dynamic, as in our simulated example. In line with this, note that below we show that our results are relatively insensitive to the value of this parameter.

@@

\begin{figure*}%[tbhp]
\centering
\includegraphics[width=0.80\linewidth]{figures/Figure2/Figure2_v5}
\caption{The model can decode hippocampal replay trajectories using either sorted and clusterless spikes from the same SWR event. \textbf{(A)} Decoding using sorted spikes. The top panel shows 31 cells on a W-track ordered according to linearized position by their place field peaks. The middle panel shows the probability of each dynamic over time as in Figure \ref{1}F. Shaded regions correspond to the speed classifications as in Figure \ref{1}G. The bottom panel shows the estimated probability of latent position over the course of the SWR as it travels down the center arm toward the right arm. L, R, C correspond to the position of the left, right and center reward wells respectively. The animal's actual position is indicated by the the magenta dashed line. Underneath is the maximum of the 1D decoded position (the most probable position) projected back onto the 2D track for the sorted decoding. Color indicates time. The animal's actual position is denoted by the pink dot. Light grey lines show the animal's 2D position over the entire recording session. \textbf{(B)} Decoding using clusterless spikes. The top panel shows multiunit spiking activity from each tetrode. Other panels have the same convention as (A). Underneath is the maximum of the 1D decoded position (the most probable position) projected back into 2D using the clusterless decoding.
}
\label{2}
\end{figure*}
\subsection*{Identification of continuous dynamics in a SWR with sorted or clusterless spikes}

@@

We next sought to validate the model on real hippocampal data recorded from awake, behaving rats. To ensure that we could capture rapid trajectories many times the speed of the animal---as described by most studies of hippocampal replay---we first decoded a single SWR with sequential population activity (Figure \ref{2}A, top panel) from 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on a W-shaped track (see Methods for details). As expected, we observed that the probability of being in the continuous dynamic is high throughout this event, but the probability of being in a stationary dynamic was also noticeable at the beginning and end of the SWR (Figure \ref{2}A, middle panel). Using our speed classification scheme, this means that the speed of the replay starts slower---as a mixture of continuous and stationary dynamics---and then speeds up and slows down again. This is also evident in the posterior probability of latent linear position over time. This shows that the replay most likely travels down the center arm and up the right arm (Figure \ref{2}A, bottom panel). We can also see this when we project the maximum of the posterior of this trajectory (the most probable "mental" position) to 2D to better see the spatial trajectory on the maze (Figure \ref{2}C). Importantly, when we apply the same model using the 2D position of the animal, we get a similar result as when we use the 1D linearized position (Figure \ref{fig:Figure2-Figure supplement 1}A).

One of our criteria for a more optimal method is for it to use all of the available spiking data. Using only clustered spikes discards any spike events that cannot be uniquely assigned to a putative single neuron, substantially reducing the amount of information that the resultant decoding can use. Additionally, spike sorting is not necessary to recover the underlying neural population dynamics \cite{TrautmannAccurateEstimationNeural2019}, is time-consuming, and often requires inclusion decisions that can vary widely across experimenters. We therefore adopted a "clusterless" approach which directly relates spikes and their multiunit spike waveform features to decode position without spike sorting (see Methods). Clusterless decoding has previously been used to successfully identify theta sequences and replay sequences in the hippocampus \cite{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}. Applying a clusterless decoder to the same SWR event, we get similar classification of the sequence (Figure \ref{2}B, D), both with 1D linearized position and 2D position (Figure \ref{fig:Figure2-Figure supplement 1}B). As predicted, the spatial extent of the event is longer and the estimate of the posterior probability of latent position is narrower for the clusterless model. This reflects the clusterless model's access to a larger pool of data that provides more information about the extent of the event and more certainty in the latent position and the dynamic (Figure \ref{2}D vs C).

@@

As expected, and providing further support for the validity of the model, we observe many replays that are classified as continuous throughout the entirety of the SWR, similar to those seen using the standard decoder (Figure \ref{fig:Figure2-Figure supplement 2}A-F). However, we also observe many events with spatially coherent content that do not have this structure. For example, we observe trajectories that start in one direction and reverse back to the original position (Figure \ref{3}A, Figure \ref{fig:Figure3-Figure supplement 1}B), trajectories that remain fixed in one position (Figure \ref{3}B, Figure \ref{fig:Figure3-Figure supplement 1}G), and trajectories that jump between arms and between dynamics (Figure \ref{3}C, D, F, Figure \ref{fig:Figure3-Figure supplement 1}F, H, I). We also observe SWRs that are spatially incoherent throughout the SWR (Figure \ref{fig:Figure3-Figure supplement 1}A, D).

@@

Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content \cite{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}. We therefore asked what fraction of classified events contain spatially coherent content, which we define as SWRs containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods and Figure \ref{1}F). We find that 96\% (22,478 of 23,382) of classified SWRs and 85\% of all SWRs include spatially coherent structure and that this prevalence of spatially coherent structure is consistent across animals (Figure \ref{3}G). We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented or fragmented-continuous mixture dynamics. We find that only 14\% (3,353 of 23,382) of classified SWRs have time points with spatially incoherent structure (Figure \ref{3}G).

@@

Because our model is specified in the context of a latent position associated with different movement dynamics, it allows us to not only classify events in terms of their dynamics, but also to quantify the model's certainty in each position estimate at each moment in time given the model parameters. To do so, we can compute the cumulative spatial size of the 95\% highest posterior density (HPD) region of the latent position estimate. Larger values of the HPD region size indicate the model is less certain about position, because the most probable decoded positions are distributed over more of the track at a given time point. In contrast, smaller values indicate that the model is more certain about the estimate of position because the extent of the HPD is more concentrated and covers less of the track. Thus, the HPD region provides a complementary measure of spatial coherence, and evaluating it allows us to verify that the events we defined as spatially coherent also correspond to events where there is high certainty around the position estimates.

We find that spatially coherent events indeed have smaller HPD regions than events with fragmented dynamics. Figure \ref{4}A and Figure \ref{4}D show two example SWRs that are classified as having stationary and continuous dynamics, respectively. The most probable positions (as estimated by the model) at each time step in these SWRs is concentrated in a small portion of the track and correspondingly, the HPD region is small throughout the SWRs. In contrast, Figure \ref{4}C shows a SWR where the dynamics are fragmented and correspondingly, the most probable positions are more spatially diffuse and the HPD region is much larger. The HPD region size also provides insights into the moment-by-moment structure of each event, which can change over the time course of a SWR. An example of this change is shown in Figure \ref{4}B, where the HPD region is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation. Overall, when we examine the average HPD region size for each SWR, grouped by dynamic, we find a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E). For the spatially coherent dynamics, the average HPD region for each SWR was much smaller than the spatially incoherent dynamics (median 24 cm, spatially coherent vs. median 238 cm, spatially incoherent, p=2.2e-16, one-sided Mann-Whitney-U). Additionally, the HPD region for the unclassified SWRs was similarly large. We note here that while the size of the HPD region will be influenced by the dynamic estimated from the data, it remains possible for continuous or stationary dynamics to correspond to a large HPD region (see outliers in Figure \ref{4}E), indicating less spatial certainty for those specific events. In general, if the data does not exhibit strong place specific firing, the HPD region will be large, regardless of dynamic classification. To show this, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model. The shuffled decodes have much larger HPD regions than the real data (epoch \#1: 55 cm vs. 229 cm, p=0.02, real vs. shuffled; epoch \#2: 72 cm vs. 231 cm, p=0.02, real vs. shuffled; one-sided Mann-Whitney-U, \ref{fig:Figure4-Figure supplement 1}).

\begin{figure*}%[tb2]
\centering
\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v6}
\caption{Validation of classification using the 95\% Highest Posterior Density. \textbf{(A-D)}  Examples of the 95 \% Highest Posterior Density. In each column: top panel: Probability of dynamic over time. Shading and labels indicate dynamic categories. Middle panel: Posterior distribution of estimated linear position over time. Magenta dashed line indicates animal's position. Bottom panel: HPD region size---the integral of the position bins with the top 95\% probable values. \textbf{(E)} Average 95\% HPD region size for each dynamic category. Median 24 cm for spatially coherent vs. median 238 cm for spatially incoherent, p=2.2e-16, one-sided Mann-Whitney-U}
\label{4}
\end{figure*}

\subsection*{Many replay events are slow or stationary}
\begin{figure*}%[tbhp]
\centering
\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v4}
\caption{
Prevalence of classifications. \textbf{(A)} UpSet plot \cite{LexUpSetVisualizationIntersecting2014}---which is similar to a Venn diagram with more than 3 sets---of the most common sets of classifications within each SWR. Each row represents a classification and each column represents a set of classifications, where filled-in black dots with an edge between the dots indicates that multiple classifications are present in the SWR (at temporally distinct times). The sets are ordered by how often they occur as indicated by the bar plot above each category. The total percentage of each classification is indicated by the rotated bar plot on the left. \textbf{(B)} Duration of each dynamic category within a SWR. The box shows the interquartile range (25-75\%) of the data and the whiskers are 1.5 times the interquartile range. Outlier points outside of this are labeled as diamonds. \textbf{(C)} Average distance between the latent position and the animal's actual position for each classification within the SWR. \textbf{(D)} Average speed of the classification within the SWR (excluding classifications with durations < 20 ms). Note that these speeds are calculated using the most probable position (MAP estimate) which can be noisy when the probability of position is flat or multimodal.\textbf{(E)} Kernel density estimate of the position of stationary trajectories on the W-track at least 30 cm away from the animal's position. The shaded region represents the density estimate while the black ticks represent the observed non-local stationary positions. \textbf{(F)} Average tetrode multiunit spike rates for each dynamic category within each SWR (excluding classifications < 20 ms).
}
\label{5}
\end{figure*}

@@

Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of <1 m/s. The next most common category were stationary events \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events. Importantly, events with slower dynamics tended to persist for approximately the same length of time as events (\texttildelow50-100 ms) with faster continuous dynamics, consistent with a brief activation of a representation of a location or a short trajectory through space. Events with slower dynamics could also represent either the animal's current location or locations far from the animal, and similarly, previous work has shown that events with continuous dynamics most often begin at the animal's current location \cite{DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.

@@

\subsection*{Highest Posterior Density}
The 95\% highest posterior density (HPD) is a measurement of the spread of the posterior probability and is defined as the region of the posterior that contains the top 95\% of values of the posterior probability \cite{CasellaStatisticalinference2001}. In this manuscript, we use the HPD size to evaluate the uncertainty of the posterior probability of position. We calculate the 95\% HPD region by determining the maximum threshold value $h$ that fulfills the follow equality:
$$
\int_{\{x: p(x_{k} \mid O_{1:T}) > h\}} p(x_{k} \mid O_{1:T})dx = 0.95
$$
The 95\% HPD region is the set of position bins with posterior values greater than the threshold $\{x : p(x_{k} \mid O_{1:T}) > h\}$. The HPD size is calculated by taking the integral of the members of this set:
$$
\int_{\{x: p(x_{k} \mid O_{1:T}) > h\}} \mathbb{1}dx
$$
which we approximate with a Riemann sum.
